---
title: "R Notebook"
output: html_notebook
---



```{r}
library(tidyverse)
```

```{r}
library(tools)
```

```{r}
library(janitor)
```

```{r}
decathlon <-
read_rds("dirty_data_project_raw_data/decathlon.rds")
```

#just having a look at the data
```{r}
decathlon %>% 
  names()
```

#placing athlete names in new column, away from inherent row name
```{r}
decathlon <- decathlon %>% 
  rownames_to_column(var = "name")
```

#seeing if there are any NA values to worry about. None.
```{r}
sum(is.na(decathlon))
```

#getting a list of all of the column names in one place, so i don't have to refer back to a table for the rearranging that's about to be done.
#rearranging columns for a more organised reading. Basic details of competition, rank, name, and points first. Followed by the runs. Followed by jumps. Followed by throws.

```{r}
decathlon %>% 
  colnames()
```

```{r}
decathlon_rearranged <- decathlon[c("Competition", "Rank", "name", "Points", "100m", "400m", "1500m", "110m.hurdle", "Long.jump", "High.jump", "Pole.vault" , "Javeline", "Shot.put", "Discus" )]
```


```{r}
decathlon_rearranged <- clean_names(decathlon_rearranged)
decathlon_rearranged
```


#changing the athlete names from all-caps to title. Assigning to new dataframe because I think the data is looking clean now.
```{r}
decathlon_clean <-
decathlon_rearranged %>% 
  mutate(name = str_to_title(name))
decathlon_clean
```
#1.1.2.1

#checking the number for highest long jump number
```{r}
max(decathlon_clean$long_jump)
```

#pulling up all details of the highest long jump. Matches with number above.
```{r}
arrange(decathlon_clean, desc(long_jump)) %>% 
  head(1)
```

#1.1.2.2

#summarising the average of the competition column when it's been filtered down to each competition
```{r}
decathlon_clean %>% 
  summarise(average_100_Decastar = mean(filter(decathlon_clean, competition == "Decastar")
                                        $x100m, na.rm = TRUE)
            )
decathlon_clean %>% 
  summarise(average_100_OlympicG = mean(filter(decathlon_clean, competition == "OlympicG")
                                        $x100m, na.rm = TRUE)
            )
```


#1.1.2.3

```{r}
decathlon_clean %>% 
  group_by(name) %>% 
  summarise(overall_points = (sum(points))) %>% 
  arrange(desc(overall_points)) %>% 
  head(1)
```



#1.1.2.4

```{r}
decathlon_clean %>% 
  select(competition, name, rank, shot_put) %>% 
  filter(competition == "Decastar") %>%
  arrange(desc(shot_put)) %>% 
  head(3)

decathlon_clean %>% 
  select(competition, name, rank, shot_put) %>% 
  filter(competition == "OlympicG") %>%
  arrange(desc(shot_put)) %>% 
  head(3)

```

#1.1.2.5
```{r}
decathlon_clean %>% 
  summarise(more_than_50 = 
              mean(
                filter(
                  decathlon_clean, x400m >= 50)$x400m, na.rm = TRUE))

decathlon_clean %>% 
  summarise(less_than_50 = 
              mean(
                filter(
                  decathlon_clean, x400m <= 50)$x400m, na.rm = TRUE))


```




  summarise(average_100 = mean(filter(decathlon_clean, competition == "Decastar")$"100m", na.rm = TRUE))





```{r}

```






























































































































































































































































































































































