---
title: "R Notebook"
output: html_notebook
---



```{r}
library(tidyverse)
```

```{r}
decathlon <-
read_rds("dirty_data_project_raw_data/decathlon.rds")
```



```{r}
decathlon
```



```{r}
dog <-
read_csv("dirty_data_project_raw_data/dog_survey.csv")
```


```{r}
decathlon %>% 
  names()
```


```{r}
decathlon <- decathlon %>% 
  rownames_to_column(var = "name")
```


```{r}
sum(is.na(decathlon))
```


```{r}
decathlon
```
```{r}
decathlon %>% 
colnames()
```

```{r}
decathlon_rearranged <- decathlon[c("Competition", "Rank", "name", "Points", "100m", "400m", "1500m", "110m.hurdle", "Long.jump", "High.jump", "Pole.vault" , "Javeline", "Shot.put", "Discus" )]
```


```{r}
decathlon_rearranged
```
```{r}
library(tools)
```


```{r}
decathlon_rearranged_case$name <- tolower(decathlon_rearranged[["name"]])
```


```{r}
decathlon_rearranged_case<-
decathlon_rearranged %>% 
  mutate(name = str_to_title(name))
```


```{r}
decathlon_rearranged_case
```

```{r}
library(snakecase)
```

decathlon_rearranged <- decathlon[c("Competition", "Rank", "name", "Points", "100m", "400m", "1500m", "110m.hurdle", "Long.jump", "High.jump", "Pole.vault" , "Javeline", "Shot.put", "Discus" )]



```{r}
decathlon_rearranged_snakecase <- 
  decathlon_rearranged_case %>% 
  rename(
    competition = Competition,
    rank = Rank,
    points = Points,
    "110m_hurdle" = "110m.hurdle",
    long_jump = Long.jump,
    high_jump = High.jump,
    pole_vault = Pole.vault,
    javeline = Javeline,
    shot_put = Shot.put,
    discus = Discus
    )
decathlon_rearranged_snakecase
```


```{r}
max(decathlon_clean$long_jump)
```



```{r}
arrange(decathlon_clean, desc(long_jump)) %>% 
  head(1)
```



```{r}
decathlon_clean %>% 
  summarise(average_100 = mean
            (filter(decathlon_clean, competition == "Decastar")
              $"100m", na.rm = TRUE))
```

```{r}
decathlon_clean %>% 
  summarise(average_100_Decastar = mean(filter(decathlon_clean, competition == "Decastar")
                                        $"100m", na.rm = TRUE)
            )
decathlon_clean %>% 
  summarise(average_100_OlympicG = mean(filter(decathlon_clean, competition == "OlympicG")
                                        $"100m", na.rm = TRUE)
            )
```



```{r}
decathlon_clean %>% 
  group_by(name) %>% 
  summarise(overall_points = (sum(points))) %>% 
  arrange(desc(overall_points))
```




```{r}
decathlon_clean %>% 
  select(competition, name, rank, shot_put) %>% 
  filter(competition == "Decastar") %>%
  arrange(desc(shot_put)) %>% 
  head(3)

decathlon_clean %>% 
  select(competition, name, rank, shot_put) %>% 
  filter(competition == "OlympicG") %>%
  arrange(desc(shot_put)) %>% 
  head(3)

```

```{r}
decathlon_clean %>% 
  mean(filter("400m" <= 50)$400m, na.rm = TRUE)
```

```{r}
library(janitor)
```


```{r}
decathlon_clean <- clean_names(decathlon_clean)
```

```{r}
decathlon_clean %>% 
  summarise(more_than_50 = 
              mean(
                filter(
                  decathlon_clean, x400m >= 50)$x400m, na.rm = TRUE))

decathlon_clean %>% 
  summarise(less_than_50 = 
              mean(
                filter(
                  decathlon_clean, x400m <= 50)$x400m, na.rm = TRUE))


```




  summarise(average_100 = mean(filter(decathlon_clean, competition == "Decastar")$"100m", na.rm = TRUE))





```{r}

```






























































































































































































































































































































































