---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}
library(readxl)
```

```{r}
library(janitor)
```

```{r}
library(tidyr)
```



#reading in the three DFs
```{r}
candy_2015 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2015.xlsx")
candy_2016 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2016.xlsx")
candy_2017 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2017.xlsx")
```

#cleaning the names and assigning to new variables
```{r}
candy_2015 <- clean_names(candy_2015)
candy_2016 <- clean_names(candy_2016)
candy_2017 <- clean_names(candy_2017)
```



#--------------------------

#giving id to each DF, but maybe not needed?

```{r}
candy_2015 <-
candy_2015 %>% 
  mutate(id = 1:5630) %>% 
  relocate("id", .before = 1)
```
#-------------------



#renaming the age value so there's something to bind all three dfs to

```{r}
candy_2015 <-
candy_2015 %>% 
  rename(age = how_old_are_you)
```

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(age = how_old_are_you)
```

```{r}
candy_2017 <-
candy_2017 %>% 
  rename(age = q3_age)
```


#renaming the going out columns in 15/16


```{r}
candy_2015 <-
candy_2015 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself)
```

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself)
```

#comparing the column naems of dfs. really helpful to see what could be renamed to match
```{r}
compare_df_cols(candy_2015, candy_2016, candy_2017)
```


#renaming mary janes
```{r}
candy_2015 <-
candy_2015 %>% 
rename(anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes = anonymous_brown_globs_that_come_in_black_and_orange_wrappers)
```

```{r}
candy_2016 <-
candy_2016 %>% 
rename(anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes = anonymous_brown_globs_that_come_in_black_and_orange_wrappers)
```

#renaming boxo raisins
_________
```{r
candy_2017 <-
candy_2017 %>% 
  rename(box_o_raisins = boxo_raisins)
```
#mixed up after having to rerun chunks. Is above code necessary?

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(box_o_raisins = boxo_raisins)
```

#renaming country and state


#comparing the column naems of dfs. really helpful to see what could be renamed to match
```{r}
compare_df_cols(candy_2015, candy_2016, candy_2017)
```

```{r}
candy_bind <-
  bind_rows(candy_2015, candy_2016, candy_2017)
```




#How I'm going to clean the data:
#1. Parse anything from the datasets that isn't pertinent to the subsequent analysis. This will involve parsing out the Degrees of Separation questions and the other random question
#2. Pivot the DFs to longform, so that there is only one column for candy and one for the participant's rating
#3. Change the column names that have corresponding data in other DFs, and clean up the values within them. Specifically Countries and States.
#4. Merge the three DFs, 













1.4 Task 4 - Halloween Candy Data
The data is in files boing-boing-candy-2015.xlxs, boing-boing-candy-2016.xlxs and boing-boing-candy-2017.xlxs. Bear in mind that this is trickier compared with tasks 1, 2 & 3.

More information on this data is available here

1.4.1 Some cleaning hints

You’ll need to combine these three datasets together.
The column country is particularly messy, you will likely need to do some ‘hard coding’ here!

1.4.2 Analysis questions

What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating?
For each of joy, despair and meh, which candy bar revived the most of these ratings?
How many people rated Starburst as despair?
For the next three questions, count despair as -1, joy as +1 and meh as 0.

What was the most popular candy bar by this rating system for each gender in the dataset?
What was the most popular candy bar in each year?
What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?













#binding all three dfs for the first time
#---------------
```{r}
bind_candy <- bind_rows(candy_2015, candy_2016, candy_2017)
```
#---------------



```{r}
candy_2016 %>% 
  names()
```




```{r}
candy_2017 %>% 
  names()
```


```{r}
candy_2017
```


```{r}
names(candy_2017) <- sub("q1_", "", names(candy_2017))
names(candy_2017) <- sub("q2_", "", names(candy_2017))
names(candy_2017) <- sub("q3_", "", names(candy_2017))
names(candy_2017) <- sub("q4_", "", names(candy_2017))
names(candy_2017) <- sub("q5_", "", names(candy_2017))
names(candy_2017) <- sub("q6_", "", names(candy_2017))
names(candy_2017) <- sub("q7_", "", names(candy_2017))
names(candy_2017) <- sub("q8_", "", names(candy_2017))
names(candy_2017) <- sub("q9_", "", names(candy_2017))
names(candy_2017) <- sub("q10_", "", names(candy_2017))
names(candy_2017) <- sub("q11_", "", names(candy_2017))
names(candy_2017) <- sub("q12_", "", names(candy_2017))
names(candy_2017) <- sub("q13_", "", names(candy_2017))

```

```{r}
candy_2017 %>% 
  str_replace_all(sub
```
#figure out later








```{r}
candy_2017 %>% 
  names()
```
































































































































































































































