---
title: "R Notebook"
output: html_notebook
---

#How I'm going to clean the data:
#1. Parse anything from the datasets that isn't pertinent to the subsequent analysis. This will involve parsing out the Degrees of Separation questions and the other random questions. Keeping gender, age, country etc because those are asked about.
#2. Pivot the DFs to longform, so that there is only one column for candy and one for the participant's rating
#3. Change the column names that have corresponding data in other DFs, and clean up the values within them. Specifically Countries and States.
#4. Merge the three DFs.


#loading packages

```{r}
library(tidyverse)
```

```{r}
library(dplyr)
```

```{r}
library(readxl)
```

```{r}
library(janitor)
```

```{r}
library(naniar)
```


#CLEANING

#reading in the three DFs
```{r}
candy_2015 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2015.xlsx")
candy_2016 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2016.xlsx")
candy_2017 <- read_excel("dirty_data_project_raw_data/candy_ranking_data/boing-boing-candy-2017.xlsx")
```


#cleaning the names 
```{r}
candy_2015 <- clean_names(candy_2015)
candy_2016 <- clean_names(candy_2016)
candy_2017 <- clean_names(candy_2017)
```


#giving year and id columns to each to each DF

#5630 obvs for 2015
```{r}
candy_2015 <-
candy_2015 %>% 
  mutate(id = 1:5630) %>% 
  relocate("id", .before = 1) %>% 
  mutate(year = 2015)
```


#1259 obvs for 2016
```{r}
candy_2016 <-
candy_2016 %>%
  mutate(id = 5631:6889) %>% 
  relocate("id", .before = 1) %>% 
  mutate(year = 2016)
```


#2460 obvs for 2017
```{r}
candy_2017 <-
candy_2017 %>% 
  mutate(id = 6890:9349) %>% 
  relocate("id", .before = 1) %>% 
  mutate(year = 2017)
```



#renaming the age column so there's something to bind all three dfs to

```{r}
candy_2015 <-
candy_2015 %>% 
  rename(age = how_old_are_you)
```

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(age = how_old_are_you)
```

```{r}
candy_2017 <-
candy_2017 %>% 
  rename(age = q3_age)
```


#Examined DFs in relation to questions, decided on columns that are necessary for the analysis: CANDY    YEAR    AGE   GOING OUT   GENDER    COUNTRY



#renaming the going out columns in 15/16

```{r}
candy_2015 <-
candy_2015 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself)
```

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself)
```

#renaming country and state columns in 2016 and 2017

```{r}
candy_2016 <-
candy_2016 %>% 
  rename(country = which_country_do_you_live_in)
```

#deselcting some non-candy item columns from all years

#getting rid of all non-candy items

#for 2015
```{r}
candy_2015 <-
candy_2015 %>% 
  select(-cash_or_other_forms_of_legal_tender, -dental_paraphenalia, -generic_brand_acetaminophen, -glow_sticks, -broken_glow_stick, -creepy_religious_comics_chick_tracts, -healthy_fruit, -hugs_actual_physical_hugs, -kale_smoothie, -lapel_pins, -joy_joy_mit_iodine, -chick_o_sticks_we_don_t_know_what_that_is, -spotted_dick, peanut_butter_jars, -vicodin, -white_bread, -whole_wheat_anything)
```

#for 2016
```{r}
candy_2016 <-
candy_2016 %>% 
  select(-cash_or_other_forms_of_legal_tender, -dental_paraphenalia, -generic_brand_acetaminophen, -glow_sticks, -broken_glow_stick, -creepy_religious_comics_chick_tracts, -healthy_fruit, -hugs_actual_physical_hugs, -kale_smoothie, -joy_joy_mit_iodine, -chick_o_sticks_we_don_t_know_what_that_is, -spotted_dick, -vicodin, -white_bread, -whole_wheat_anything, -minibags_of_chips, -person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes, -chardonnay, -bonkers_the_board_game)

```

#for 2017

#removing all of the unnecessary Qs
```{r}
names(candy_2017) <- sub("q[0-9]_", "", names(candy_2017))
names(candy_2017) <- sub("q1[0_9]_", "", names(candy_2017))
```

#renaming this one because it had a "Q6" in front of it. Better to have column names that don't begin with a letter.
```{r}
candy_2017 <-
candy_2017 %>% 
  rename(x100_grand_bar = "100_grand_bar")
```

#and the removing non-candy items, just like for other two years
```{r}
candy_2017 <-
candy_2017 %>% 
  select(-cash_or_other_forms_of_legal_tender, -dental_paraphenalia, -generic_brand_acetaminophen, -glow_sticks, -broken_glow_stick, -creepy_religious_comics_chick_tracts, -healthy_fruit, -hugs_actual_physical_hugs, -kale_smoothie, -joy_joy_mit_iodine, -chick_o_sticks_we_don_t_know_what_that_is, -spotted_dick, -vicodin, -white_bread, -whole_wheat_anything, -minibags_of_chips, -chardonnay, -bonkers_the_board_game, -real_housewives_of_orange_county_season_9_blue_ray)
```


#turning candy into observations, instead of each candy having one column each. Using pivot longer to do this.


#pivoting 2015, rearranging position of columns
```{r}
candy_2015_long <-
candy_2015 %>% 
  pivot_longer(cols = c("x100_grand_bar":"york_peppermint_patties", butterfinger, sea_salt_flavored_stuff_probably_chocolate_since_this_is_the_it_flavor_of_the_year, necco_wafers),
               names_to = "candy",
               values_to = "rating")
```

#choosing the columns from before that I figured were necessary for the analysis.
```{r}
candy_2015_long <-
candy_2015_long %>% 
  select("id":"going_out", candy, rating, year) %>% 
  relocate(year, id, age, candy, rating, going_out, timestamp)
candy_2015_long
```
#2015 looking ok for now. Onto 2016

#same process, pivot longer
```{r}
candy_2016_long <-
candy_2016 %>% 
  pivot_longer(cols = c("x100_grand_bar":"york_peppermint_patties"),
               names_to = "candy",
               values_to = "rating")

```


#and then choose necessary columns
```{r}
candy_2016_long <-
candy_2016_long %>% 
  select("id":"state_province_county_etc", candy, rating, year) %>% 
  relocate(year, id, age, your_gender, country, state_province_county_etc, candy, rating, going_out, timestamp)
```

#2016 looking ok. Need to figure out a way to assign specific items to JOY and DESPAIR columns. ASK INSTRUCTORS TMR
#_______________________________________________________________________________-

#and same for 2017 as other two
```{r}
candy_2017_long <-
candy_2017 %>% 
  pivot_longer(cols = c("x100_grand_bar":"york_peppermint_patties"),
               names_to = "candy",
               values_to = "rating")

```


```{r}
candy_2017_long <-
candy_2017_long %>% 
  select("id":"state_province_county_etc", candy, rating, year) %>% 
  relocate(year, id, age, gender, country, state_province_county_etc, candy, rating, going_out)
```

#comparing columns of the three dfs, seeing if they look ready to bind into one.
```{r}
compare_df_cols(candy_2015_long, candy_2016_long, candy_2017_long)
```

#all three DFs look ok. Time to bind them.
```{r}
candy_bind <-
  bind_rows(candy_2015_long, candy_2016_long, candy_2017_long)
candy_bind
```



#get rid of rows that have no rating value and no going out value

#originally tried this line of code but it cut out too much, removed all rows which had an NA value in going_out or rating. Logically the statement reads "show results if it is NOT the case that going_out is NA, and filter results if it is NOT the case that rating is NA. Many rows could have an NA value in one but not the other, and still get filtered out. This isn't what I want.
```{r}
#candy_bind %>% 
#  filter(!is.na(going_out) & !is.na(rating))
```

#this one reads "show results if it is NOT the case that BOTH going_out and rating are NA. There are fewer results where this is the case. 
```{r}
candy_bind_filtered <-
candy_bind %>% 
  filter(!(is.na(going_out) & is.na(rating)))
candy_bind_filtered
```



#QUESTIONS


#1.4.2.1
#What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candy_bind_filtered %>%
  count(NA_value_for_rating = is.na(rating))

candy_bind_filtered %>% 
count(rating)
  
```
#the code has returned FALSE to is.na for 637584 values. Second bit of the code shows that there are only three non-NA values, of DESPAIR, JOY, and MEH, so no other weird non-answers.

#____________________________________________________________





#1.4.2.2


#First I'm extracting only numeric values from the age column- there were a couple of values in there that looked really messy, a jumble of letters and numbers. Then I'm changing the age column type to numeric.

#In the second part only age and going_out columns are selected, because they're the ones up for analysis. Then, filtering out the ridiculous numbers, an age range has been set from 3 (when it might be reasonable to start answering telephones) to 122 (the oldest living person). Arrange, group, summarise mean.

```{r}
  candy_bind_filtered_age <-
  candy_bind_filtered %>%
  mutate(age = str_extract(age, "^[0-9]*+")) %>%
  mutate(age =  as.numeric(age))

  candy_bind_filtered_age %>% 
  select(age, going_out) %>% 
  filter(age <= 122 & age >= 3) %>% 
  arrange(desc(age)) %>% 
  group_by(going_out) %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```


#1.4.2.3

#ordered and then assigned to new variable. Filtered out each rating value within that new variable, because there were only three.


```{r}
candy_bind_rating_count <-
candy_bind_filtered_age %>% 
  group_by(candy) %>% 
  count(rating) %>% 
  arrange(desc(n))
  
  candy_bind_rating_count %>% 
  filter(rating == "JOY") %>% 
    head(1)

  candy_bind_rating_count %>% 
  filter(rating == "DESPAIR") %>% 
    head(1)

  candy_bind_rating_count %>% 
  filter(rating == "MEH") %>% 
    head(1)

```


#1.4.2.4

```{r}
candy_bind_filtered_age %>% 
  group_by(candy) %>% 
  count(rating) %>% 
  filter(candy == "starburst") %>% 
  filter(rating == "DESPAIR")
```

#1.4.2.6

#First, I'm using case_when to create a new column where each of the ratings has a numeric equivalent. Using the same filter for age (between 3 and 122). Assigning all of this to a new variable for shorthand.

```{r}
candy_bind_rating_points <-
candy_bind_filtered_age_ex %>% 
mutate(rating_points = case_when(
  rating == "DESPAIR" ~ -1,
  rating == "JOY" ~ 1,
  rating == "MEH" ~ 0
  )) %>% 
  filter(age <= 122 & age >= 3)
candy_bind_rating_points
```


#Then using new variable to run 4 variants of the same code chunk, for each of the 4 gender responses that wasn't NA. I'm filtering out by the gender response, grouping by candy, creating a new column with the sumrs of the ratings and taking the top value. 

#for all four chunks the result shows that any full sized candy bar is the most highly rated candy bar. In second place, male and female respondents both rated reese's peanut butter cups most highly, but those who didn't leave their gender rated kit kats and those with gender "other" rated twix.

```{r}
candy_bind_rating_points %>% 
  filter(gender == "Male") %>% 
  group_by(candy) %>% 
  summarise(candy_points_male = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_male)) %>% 
  head(2)

candy_bind_rating_points %>% 
  filter(gender == "Female") %>% 
  group_by(candy) %>% 
  summarise(candy_points_female = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_female)) %>% 
  head(2)

candy_bind_rating_points %>% 
  filter(gender == "I'd rather not say") %>% 
  group_by(candy) %>% 
  summarise(candy_points_rather_not_say = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_rather_not_say)) %>% 
  head(2)

candy_bind_rating_points %>% 
  filter(gender == "Other") %>% 
  group_by(candy) %>% 
  summarise(candy_points_other = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_other)) %>% 
  head(2)
```




#1.4.2.7

#effectively reworked the same chunk of code for question 6, except for year. Once again, any full sized candy bar was the most highly rated across the board.

```{r}
candy_bind_rating_points %>% 
  filter(year == 2015) %>% 
  group_by(candy) %>% 
  summarise(candy_points_2015 = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_2015)) %>% 
  head(1)

candy_bind_rating_points %>% 
  filter(year == 2016) %>% 
  group_by(candy) %>% 
  summarise(candy_points_2016 = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_2016)) %>% 
  head(1)

candy_bind_rating_points %>% 
  filter(year == 2017) %>% 
  group_by(candy) %>% 
  summarise(candy_points_2017 = (sum(rating_points, na.rm = TRUE))) %>% 
  arrange(desc(candy_points_2017)) %>% 
  head(1)

```


#various attempts at question 8 are in halloween_candy, line 272.








